---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# quanteda.proximity

<!-- badges: start -->
[![R-CMD-check](https://github.com/gesistsa/quanteda.proximity/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/gesistsa/quanteda.proximity/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The goal of quanteda.proximity is to add a hacky layer of proximityp vectors into the `tokens` object of `quanteda`.

## Installation

You can install the development version of quanteda.proximity like so:

``` r
remotes::install_github("gesistsa/quanteda.proximity")
```

## Example

```{r example}
suppressPackageStartupMessages(library(quanteda))
library(quanteda.proximity)

testdata <-
c("Turkish President Tayyip Erdogan, in his strongest comments yet on the Gaza conflict, said on Wednesday the Palestinian militant group Hamas was not a terrorist organisation but a liberation group fighting to protect Palestinian lands.",
"EU policymakers proposed the new agency in 2021 to stop financial firms from aiding criminals and terrorists. Brussels has so far relied on national regulators with no EU authority to stop money laundering and terrorist financing running into billions of euros.")
```

`tokens_proximity()` generates the proximity vectors and stores them as a `docvar` (document variable).

```{r tokens_proximity}
res <- testdata %>% tokens() %>% tokens_tolower() %>%
    tokens_proximity(keywords = "turkish")
res
```

You can access the proximity vectors by

```{r proximity_vectors}
docvars(res, "proximity")
```

The `tokens` object with proximity vectors can be converted to a (weighted) `dfm` (Document-Feature Matrix). The default weight is assigned by inverting the proximity.

```{r dfm}
dfm(res)
```

You have the freedom to change to another weight function. For example, not inverting.

```{r dfm2}
dfm(res, weight_function = identity)
```

Or any custom function

```{r dfm3}
dfm(res, weight_function = function(x) { 1 / x^2 })
```

## Application

A clumsy example to calculate the total inverse proximity weighted frequency of "terror*" words.

```{r}
terror_dict <- dictionary(list(TERROR = c("terror*")))

dfm(res) %>% dfm_lookup(terror_dict) %>% rowSums()
```

How about changing the target to "Hamas"?

```{r}
res2 <- res %>% tokens_proximity(keywords = "hamas")
res2
```

```{r}
dfm(res2) %>% dfm_lookup(terror_dict) %>% rowSums()
```

Can we use two targets, e.g. "EU" and "Brussels"?

```{r}
res3 <- res %>% tokens_proximity(keywords = c("eu", "brussels"))
res3
```

```{r}
docvars(res3, "proximity")
```

```{r}
dfm(res3) %>% dfm_lookup(terror_dict) %>% rowSums()
```
